---
title: "diamonds"
author: "James Hamilton"
date: "March 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
data(diamonds)
```

### Histogram all diamond prices - used log10 scale for X axis to get rid of long tail.
```{r}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(binwidth=.01) +
  scale_x_log10()
```

### Check out the big peak (first peak) by limiting the x axis values
```{r}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(limits = c(300,1500))
```

### Generate a faceted histogram of prices across 'cuts'
```{r}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(binwidth = 100) +
  facet_wrap(~cut)

by(diamonds$price, diamonds$cut, max)

```

### New faceted histogram with "free" scales on Y axis
```{r}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(binwidth = 100) +
  facet_wrap(~cut, scales = "free_y")
```

### Faceted price per carat by cut
```{r}
ggplot(diamonds, aes(x = price / carat)) +
  geom_histogram(bins=1000) +
  scale_x_log10() +
  facet_wrap(~cut, scales = "free_y")
```

### Box plot and summary analysis of price by clarity (colors!)
```{r}
by(diamonds$price, diamonds$clarity, summary)

ggplot(diamonds, aes(x = clarity, y = price)) +
  geom_boxplot(aes(fill=clarity),color="black") +
  coord_cartesian(ylim = c(300, 6300)) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank())
```

### Examine IQR across colors
```{r}
diamonds %>%
  group_by(color) %>%
  summarise(low = min(price),
    '25p' = quantile(price, probs=0.25),
    mid = median(price),
    avg = mean(price),
    '75p' = quantile(price, probs=0.75),
    high = max(price),
    interqr = IQR(price)) %>%
  arrange(color)
```

###  price per carat by color boxplots
```{r}
ggplot(diamonds, aes(x = color, y = price/carat)) +
  geom_boxplot(aes(fill=color),color="black") +
  coord_cartesian(ylim = c(300, 6300)) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank())
```

### carat investigation with frequency polygon
```{r}
ggplot(diamonds, aes(x = carat, y = ..count..)) +
  geom_freqpoly(binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0.1,1.25,.1), limits = c(0.1,1.25)) +
  scale_y_continuous(breaks = seq(0,5000,500))
```

